---
title: "Assignment 1"
author: "Aijing"
---


```{r, results='hide', error=FALSE, warning=FALSE,message=FALSE}
library(tidycensus)
library(tidyverse)
```

First, I want to look at what variables are in acs5 dataset:
```{r}
area_vars_2018 <- load_variables(2018, "acs5")
```


## Variables Choosen:

I want to study whether neighborhoods composed of a majority of racial minority or migrants have different commuting patterns in Suffolk Country, MA. I included median income because it is likely that income will also significantly contribute to the commuting behaviors. Eventually, I chose the following variables:

1.Median income (continuous) : 'med_income'

2.Percentage of of labour force using public transportation to workplace as percentage (continuous): 'pct_publictrans'

3.Travel time to workplace for labor force (continuous): 'time_travel_work'

4.Majority race of the population (not Hispanic or Latino) (categorical) : 'maj_race'

5.Whether the majority of the population is foreign born or not(categorical): 'maj_foreign'


```{r}
pop_vars <- c(
          tot_pop = "B01003_001",
          popln_inwork_16 = "B23025_002",
          popln_25_64 = "B23006_001",
          med_income = "B06011_001",
          popln_publictrans_work = "B08119_028",
          time_travel_work = "B08013_001", 
          white =       "B03002_003",
          black_AA =    "B03002_004",
          native =      "B03002_005",
          asian =       "B03002_006",
          hawaiian = "B03002_007",
          other =             "B03002_008",
          mult_races =        "B03002_009",
          hispanic_latino =   "B03002_012",
          foreign_born =      "B05002_013")



tracts2018 <- get_acs(geography = "tract", 
                        variables = pop_vars, 
                        year  = 2018,
                        output = "wide",
                        state = "MA",
                        county = "Suffolk",
                        survey = "acs5")

```
## Mutate the Variables:

To obtain 'maj_race', 'maj_foreign' & 'pct_publictrans', I did the following calculations:

'pct_publictrans' is obtained by dividing larbor force('popln_inwork_16') by part of larbor force using public transportation to workplace ('popln_publictrans_work')

'maj_foreign' is obtained by evaluate whether whether foreign-born population constitutes more than 50% of the population. If so, the value will be 'TRUE
', otherwise 'FALSE'.

'maj_race' is obtained by evaluate if any race group constitues more than 50% of the population, if so that race will be designated as the majority race, otherwise 'no_majority' will be assigned.

After the first-time execution, I found there is some entries with 'NaN' or NA, so I also added a line of filter code.

```{r}
tracts2018 <- tracts2018 %>%
  mutate(pct_white = whiteE/tot_popE,
         pct_black = black_AAE/tot_popE,
         pct_hisp_lat = hispanic_latinoE/tot_popE,
         pct_native = nativeE/tot_popE,
         pct_asian = asianE/tot_popE,
         pct_hawaiian = hawaiianE/tot_popE,
         pct_mult = mult_racesE/tot_popE,
         pct_other = otherE/tot_popE,
         pct_publictrans = popln_publictrans_workE/popln_inwork_16E,
         maj_foreign = foreign_bornE / tot_popE > 0.5 )%>%
  mutate(maj_race = case_when(pct_white > 0.5 ~ "white",
                             pct_black > 0.5 ~ "black",
                             pct_hisp_lat > 0.5 ~ "hs_latino",
                             pct_native > 0.5 ~ "native_am",
                             pct_asian > 0.5 ~ "asian",
                             pct_hawaiian > 0.5 ~ "hawaiian",
                             pct_mult > 0.5 ~ "multi",
                             pct_other > 0.5 ~ "other",
                             TRUE ~ "no_majority")) %>%
  filter(pct_publictrans != "NaN", !is.na(time_travel_workE), !is.na(med_incomeE)) %>%
  select (maj_foreign, maj_race, pct_publictrans, 
          time_travel_workE, med_incomeE)
  
```

## Display & Save Results
```{r}
tracts2018
```

```{r}
write_csv(tracts2018, "tracts2018_suffolk.csv")
```

