---
title: "Assignment5"
author: "Aijing"
date: "10/5/2020"
output: html_document
---
```{r, results='hide', error=FALSE, warning=FALSE,message=FALSE}
library(tidyverse)
library(jtools)
library(interactions)
library(knitr)
library(ggplot2)
```


# Original Model

Dependent Variable: avg travel time to workplace: avg_travel_time

Independent Variables:

1. median income: med_income

2. percentage of labour force using public transit to workplace(hereafter, public transit percentage), pct_publictrans

3. majority race: maj_race

4. majority foreign or native born: maj_foreign

```{r}
tract_data <- read_csv("tracts2018_ma.csv") %>%
  mutate(maj_race = case_when(
    maj_race=="hs_latino" ~ "Hispanic/Latino", 
    maj_race=="black" ~ "Black not Hispanic/Latino",
    maj_race=="white" ~ "-White not Hispanic/Latino",
    maj_race=="asian" ~ "Asian not Hispanic/Latino",
    maj_race=="no_majority" ~ "No Majority")) %>%
  mutate(pct_publictrans = pct_publictrans*100)
```
```{r}
options(scipen = 5, digits = 4)

model1 <- lm( avg_travel_time~ med_incomeE + pct_publictrans + maj_race + maj_foreign, 
            data = tract_data)
summary(model)
```


# Log Transformation

The difference between earning $5,000 per year and $10,000 is probably more substantial than the difference between earning $50,000 per year and $55,000 per year. In the former case, income has increased by 100%, and in the latter, income has increased by 10%. In the case of income and average travel time, I think percentage increases are more relevant than absolute increases, so below i will log-tranform my median income variable.
```{r}
model2 <- lm(avg_travel_time~ log(med_incomeE) + pct_publictrans + maj_race + maj_foreign, 
            data = tract_data)

summary(model2)
```
The r-square increased by around 0.075. Log-transformed median income have a statistically significant positive correlation with average travel time, When median income doubles(one unit increase in log), the public transit percentage will also increase by 6.26 minutes.


# Converting a continuous variable to a categorical variable
I suspect the influence of median income on public transit percentage is not in a linear but step-wise fashion, i.e. once people's income have reached certain thresholds the influence on average travel time will be more or less homogeneous within that band. So I decide to test out this by divide median income variable by 4 bands: low-income 0~25,000, lower_medium-income:25,000~50,000, higher_medium-income:50,000~75,000 and high-income: 75,000 above.
```{r}
tract_data2 <- tract_data %>%
  mutate(income_group = case_when(med_incomeE < 25000 ~ "-Low_income",
                               25000 <= med_incomeE &  med_incomeE< 50000 ~ "Lower_Medium_income",
                               50000 <= med_incomeE &  med_incomeE< 75000 ~ "Higher_Medium_income",
                               med_incomeE >= 75000 ~ "High_income"))

model3 <- lm(avg_travel_time~ income_group +  pct_publictrans+ maj_race + maj_foreign, 
            data = tract_data2)

summary(model3)
```
The overall fit slightly improves as r-square increase by 0.004 compared to my original model. Interestingly, the new model shows a move from low-income to any of the other 3 groups all have a positive statically significant correlation with average travel time at a 95% confidence level. While a move from low-income to higher medium-income has the most prominent effect: it will increase travel time by 7.63 minutes.


# Adding an interaction term
I wonder if median income have a different impact on different races? To test out, I add three interactive terms as below:

```{r}

model4 <- lm(avg_travel_time ~ log(med_incomeE) + pct_publictrans + maj_race + maj_foreign + log(med_incomeE):maj_race,
            data = tract_data)

summary(model4)
```
The log-transformation on median income and an interactive term between race and log of median income increase the r-square by around 0.077. Log-transformed median income have a statistically significant extra impact on Hispanic/Latino group: every doubling in median income in a tract with majority Hispanic/Latino will increase the travel time by 4.53 minute. However, I was not able to find such a statistically significant extra impact of median income for other racial groups.


# Visualizing the results
```{r}
ModelFit <- tibble(model = c(1, 2, 3, 4),
                   R_square = c(summary(model1)$adj.r.squared,
                                summary(model2)$adj.r.squared,
                                summary(model3)$adj.r.squared,
                                summary(model4)$adj.r.squared))

ggplot(ModelFit, aes(x = model, y = R_square)) +
  geom_line() +
  scale_x_continuous(name = "",
                   breaks = breaks <- seq(1, 4, by = 1),
                   labels = paste("Model", breaks)) +
  scale_y_continuous(name = "Adjusted R-squared value") +
  theme_bw()
```

The graph tells me that model4 has the best improvement on overall fit of my regression model.

Below I will compare the coefficient change in each variable between model1 and model4 in a table.
 

```{r}
coeff_names <- c(
                 "Public transit percentage" = "pct_publictrans",
                 "Log of median Income" = "log(med_incomeE)",
                 "Median Income" = "med_incomeE",
          
                 "Asian(compared to White)" = "maj_raceAsian not Hispanic/Latino",
                 "Black(compared to White)" = "maj_raceBlack not Hispanic/Latino",
                 "Hispanic and Latino(compared to White)" = "maj_raceBlack not Hispanic/Latino",
                 "No Racial Majority(compared to White)" = "maj_raceNo Majority",
                 
                 "Majority Foreign Born(compared to Native)" = 'maj_foreignTRUE',
                 
                 "Interaction:\nAsian and Log of Median Income" = 
                    "log(med_incomeE):maj_raceAsian not Hispanic/Latino",
                 
                 "Interaction:\nBlack and Log of Median Income" = 
                    "log(med_incomeE):maj_raceBlack not Hispanic/Latino",
                 
                 "Interaction:\nHispanic/Latino and Log of Median Income" = 
                    "log(med_incomeE):maj_raceHispanic/Latino",
                 
                 "Interaction:\nNo Majority and Log of Median Income" = 
                    "log(med_incomeE):maj_raceNo Majority"
                  )
                 

               

export_summs(model1, model4, 
             error_format = "(p = {p.value})",
             error_pos = "same",
             model.names = c("Initial model", "Preferred model"),
             coefs = coeff_names)
```

```{r}
coeff_desc <- c("Majority Asian AND Doubling in Median Income" = 
                  "log(med_incomeE):maj_raceAsian not Hispanic/Latino",
                
                "Majority Black AND Doubling in Median Income" = 
                  "log(med_incomeE):maj_raceBlack not Hispanic/Latino",
                
                "Majority Hispanic/Latin AND Doubling in Median Income" = 
                  "log(med_incomeE):maj_raceHispanic/Latino",
                
                "No Majority AND Doubling in Median Income" = 
                  "log(med_incomeE):maj_raceNo Majority",
                 
                 "1 percentage increase in\n share of public transit" = "pct_publictrans",
          
                 "Asian\n(compared to White)" = "maj_raceAsian not Hispanic/Latino",
                 "Black\n(compared to White)" = "maj_raceBlack not Hispanic/Latino",
                 "Hispanic and Latino\n(compared to White)" = "maj_raceBlack not Hispanic/Latino",
                 "No Racial Majority\n(compared to White)" = "maj_raceNo Majority",
                 
                 "Majority Native\n(compared to Foreign-born)" = 'maj_foreignTRUE')
                 
                

plot_coefs(model4, 
           coefs = coeff_desc,
           plot.distributions = TRUE,
           rescale.distributions = TRUE) +
  scale_x_continuous(name = "Effect on percentage of workforce using public transit to workplace")
```



```{r}
interact_plot(model4, pred = med_incomeE, modx = maj_race, interval = TRUE) +
  scale_x_continuous(name = "Median income (thousands)",
                     limits = c(0, 110000),
                     breaks = breaks <- seq(0, 110000, by = 10000),
                     labels = paste("$", 
                                    prettyNum(breaks/1000, big.mark = "'"),
                                    sep = "")) +
  scale_y_continuous(name = "Average Travel Time to Work",
                     breaks = breaks <- seq(0, 50, by = 5)) +
  scale_color_discrete(name = "") +
  scale_linetype_discrete(name = "") +
  scale_fill_discrete(name = "") 
```

```{r}
cat_plot(model4, pred = maj_race, modx = maj_foreign , interval = TRUE) +
 scale_x_discrete(name = "Majority Race",
                    labels = c("Asian", "Black","HIspanic/Latino",
                               "No Majority","White")) +
  scale_y_continuous(name = "Public Transit Percenatge",
                     breaks = seq(0, 45, by = 5)) +
  scale_color_discrete(name = "") +
  scale_linetype_discrete(name = "") +
  scale_fill_discrete(name = "") 
```
Since I do not have an interactive term between majority race and nativity, the differencec between native and foreign-born are consistent among all racial groups.
